---
title: 'EDS Assignment 1: Cartography'
author: "Laurel Abowd"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE, warning= FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(spData)
library(spDataLarge)
library(tidyverse)
library(tmap)
library(here)
```


```{r}
# For this assignment, we will be working with the Boston dataset from the spData package.
boston <- st_read(system.file("shapes/boston_tracts.dbf", package="spData"))

boston_wgs <- boston %>% 
  st_transform(crs = 4326)
```
```{r}
# viewed data in ggplot
ggplot() +
  geom_sf(data = boston, geom = geometry, aes(fill = NOX))
```

```{r}
# Try out  tmap

tmap::tmap_mode(mode = "view")

tm_shape(boston_wgs) +
  tm_fill("NOX")

# As seen, the Boston data from spData is spatially incorrect. We found a coder who corrected the Boston data and made it available for use. They used the specific houses as points instead of tracks.
```

```{r}
# Downloaded fixed data from github 

boston_fixed <- read.csv(here("boston_fixed.csv")) %>% 
  st_as_sf(coords = c("LON", "LAT"), crs = 4267)  # the original data was in crs NAD27
```


```{r}
# Viewed data from github. Will be focusing on levels of nitrous oxide (NOX) at the housing units
tmap::tmap_mode(mode = "view")

tm_shape(boston_fixed) +
  tm_dots("NOX")
```

```{r}
# create a bbox that will go around boston, like the boston.dbf file
boston_region = st_bbox(c(xmin = -71.52311, xmax = -70.63823,
                      ymin = 42.00305, ymax = 42.67307),
                    crs = st_crs(boston_fixed)) %>%
  st_as_sfc()

```

```{r}
tmap::tmap_mode(mode = "view")


tm_shape(boston_fixed, bbox = boston_region) +
  tm_dots("NOX") 
```


```{r}
# Used package rosm to download a tile of Boston. Found bbox coordinates online.
library(rosm) 

boston_bbox = st_bbox(c(xmin = -71.154235, xmax = -70.975754,
                      ymin = 42.303574, ymax = 42.445101),
                    crs = st_crs(boston_fixed)) # used same projection as the boston fixed data
boston_map <- osm.raster(boston_bbox) # created a open street map raster the same size as the bbox 
```

```{r}
# Create static map in tmap

tmap::tmap_mode(mode = "plot")


tm_shape(boston_map) +
  tm_rgb(alpha = 0.75) +
tm_shape(boston_fixed, bbox = boston_region) +
  tm_dots("NOX", size = 0.1, palette = "RdPu", title = "NOX level", style = "cont") +
tm_scale_bar(position = c("left", "bottom")) + 
  tm_layout(title = "Boston Housing Data", 
            legend.bg.color = "white", 
            legend.position = c("right", "bottom"),
            title.size = 2,
            title.position = c("left", "top"),
            legend.frame = TRUE,
            title.bg.color = "white",
            title.bg.alpha = .3) +
  tm_graticules(alpha = .5) +
  tm_compass(position = c(.9, .3)) +
  tm_layout(frame = TRUE, frame.lwd = 5) 

# added scale bar, compass, legend, finalized visuals and layout
```

### This map shows the levels of nitrous oxide (NOX) pollution at housing units around Boston in 1978. This data is from the Boston housing data collected in 1978 by Harrison and Rubinfeld. The data location was corrected by GitHub user, namiyousef, and was downloaded from https://github.com/namiyousef/ai_hack_2021/blob/main/data/Boston_Housing_Market_Challenge/boston_fixed.csv
### The basemap was downloaded from the rosm package - open street maps: Dunnington, Dewey. 2019. rosm: Plot Raster Map Tiles from Open Street Map and Other Sources. Version 0.2.5. 
